Bootstrap learning via modular concept discovery

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Fixed multiple author citations and added shortcitations where author is obvious.

\section{Related Work}
The idea of discovering and using reusable subcomponents is part of an
old tradition in AI. For instance, it is one of the central themes in
Herb Simon's 1969 ``The Sciences of the
Artificial''~\cite{simon1996sciences}. Rendell's 1985
paper~\cite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
Induction Using Layered Information Compression: Tractable Feature
Formation in Search,'' presented within a classical search paradigm
the idea that compression provides a heuristic for constructing good
representations. Within the Inductive Logic Programming (ILP)
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
was also applied in the multitask
setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
learning functional programs and believe that the modularity and
compositionality afforded by this representation is necessary for
these ideas to be successful.

Genetic programming
(GP)~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96} has
tackled the problem of program learning and has explored the notion
that reusable functions are helpful. This work relies, however, on
stochastic local search over the space of programs, and automatically
discovering useful subroutines is seen as a helpful heuristic. More
recently, Liang et al.~\cite{liang10programs} used a stochastic
grammar over combinatory logic to induce shared structure in
multi-task program induction problems. We find this latter
representation compelling and use it here, but we see both this and
genetic programming as tackling the question of how to benefit from
shared structure in situations where local search might be successful
on its own. Such problems rely on domain knowledge -- whether in the
form of a fitness function or a likelihood function -- to provide
local structure to the search space.

@@

The basis combinators can themselves be expressed in the lambda
calculus (with definitions following directly from the equations
above). The lambda calculus has two basic operations -- application
and abstraction -- but in using the
combinatory logic we sequester uses of the abstraction operation
inside the combinators. In doing so, we have replaced the variable
binding of the $\lambda$ operator with the variable \emph{routing} of
these basis combinators. Our representation thus becomes
variable-free. See Liang et al.~\cite{liang10programs} for a more
detailed discussion of this routing interpretation.

@@

\subsection{Best-first enumeration of programs}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable expressions. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05}
\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73}. In
words, the enumeration procedure is best described by the following
recursion: every program is either a primitive combinator OR it is a
left child program applied to a right child program (that is, a left
child AND a right child).

>>>>

\section{Related Work}
The idea of discovering and using reusable subcomponents is part of an
old tradition in AI. For instance, it is one of the central themes in
Herb Simon's 1969 ``The Sciences of the
Artificial''~\shortcite{simon1996sciences}. Rendell's 1985
paper~\shortcite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
Induction Using Layered Information Compression: Tractable Feature
Formation in Search,'' presented within a classical search paradigm
the idea that compression provides a heuristic for constructing good
representations. Within the Inductive Logic Programming (ILP)
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
was also applied in the multitask
setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
learning functional programs and believe that the modularity and
compositionality afforded by this representation is necessary for
these ideas to be successful.

Genetic programming
(GP)~\cite{DBLP:conf/foga/Koza92,DBLP:conf/ices/KozaBAK96} has
tackled the problem of program learning and has explored the notion
that reusable functions are helpful. This work relies, however, on
stochastic local search over the space of programs, and automatically
discovering useful subroutines is seen as a helpful heuristic. More
recently, Liang et al.~\shortcite{liang10programs} used a stochastic
grammar over combinatory logic to induce shared structure in
multi-task program induction problems. We find this latter
representation compelling and use it here, but we see both this and
genetic programming as tackling the question of how to benefit from
shared structure in situations where local search might be successful
on its own. Such problems rely on domain knowledge -- whether in the
form of a fitness function or a likelihood function -- to provide
local structure to the search space.

@@

The basis combinators can themselves be expressed in the lambda
calculus (with definitions following directly from the equations
above). The lambda calculus has two basic operations -- application
and abstraction -- but in using the
combinatory logic we sequester uses of the abstraction operation
inside the combinators. In doing so, we have replaced the variable
binding of the $\lambda$ operator with the variable \emph{routing} of
these basis combinators. Our representation thus becomes
variable-free. See \cite{liang10programs} for a more
detailed discussion of this routing interpretation.

@@

\subsection{Best-first enumeration of programs}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable expressions. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05, DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles, DBLP:journals/cacm/Hall73}. In
words, the enumeration procedure is best described by the following
recursion: every program is either a primitive combinator OR it is a
left child program applied to a right child program (that is, a left
child AND a right child).
